---
title: "Statistical Inference Course Project 1"
author: "Ben White"
date: "26 July 2015"
output:
  html_document:
    keep_md: true
---


```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=8, fig.height=4,  warning=FALSE, message=FALSE)
```

The aim of this project is to investigate the exponential distribution and the Central Limit Theorem in R. The exponential distibution is simulated using `rexp(n, lambda)` where $n$ the number of results, and lambda is the rate parameter. The mean $\mu$ and standard deviation $\sigma$ are both 1/lambda. 

## Simulations

```{r, echo=FALSE}
library(ggplot2)
```

Set the constants

```{r}
lambda <- 0.2 
n <- 40 
noSimulations <- 1000 
```

Set the seed so that the test can be reproduced

```{r}
set.seed(1)
```

Run the 1000 simulations using rexp

```{r}
simulationResults <- matrix(data=rexp(n * noSimulations, lambda), nrow=noSimulations)
simulationMeans <- data.frame(means=rowMeans(simulationResults))
```

Plot of the Means

```{r}
g <- ggplot(data = simulationMeans, aes(x = means))  
g <- g + geom_histogram(fill = "salmon1",
                        binwidth = 0.15,
                        color = "black") 
g
```

## Question 1. Sample Mean versus Theoretical Mean

The mean $\mu$ of a exponential distribution is:

$\mu= \frac{1}{\lambda}$ 

```{r}
mu <- 1/lambda
mu
```

The mean of the sample means is as follows:

```{r}
meanOfMeans <- mean(simulationMeans$means)
meanOfMeans
```

It can be seen the theoretical mean of `r mu` and the mean of the sample means `r meanOfMeans` are close.

## Question 2. Sample Variance versus Theoretical Variance

The expected standard deviation $\sigma$ is:

$\sigma = \frac{1/\lambda}{\sqrt{n}}$ 

```{r}
sigma <- 1/lambda/sqrt(n)
sigma
```

The variance $Var$ is the standard deviation $\sigma$ squared:

$Var = \sigma^2$ 

```{r}
sigma2 <- sigma^2
sigma2
```

Calculating the standard deviation and variances of the simulations

```{r}
sdOfSimulation <- sd(simulationMeans$means)
sdOfSimulation

varOfSimulation <- var(simulationMeans$means)
varOfSimulation
```

## Q3. Show the distribution is approximately normally distributed

```{r}
g <- ggplot(data = simulationMeans,
            aes(x = means)) 
g <- g + geom_histogram(aes(y = ..density..),
                        fill = "salmon1", 
                        binwidth = 0.15,
                        color = "black")
g <- g + stat_function(fun = dnorm,
                       arg = list(mean = mu , sd = sigma),
                       colour = "blue",
                       size = 2)
g
```

As the graph shows the distribution of the sample means is approximately normally distributed.
